---
output: cynkratemplate::readme_document
---

<!-- README.md and index.md are generated from README.Rmd.
     Edit that file and render it the usual way: rmarkdown::render(), devtools::build_readme(), or the Knit button.
     The cynkratemplate package must be installed; it supplies the output format. -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# adbi

<!-- badges: start -->
[![rcc](https://github.com/r-dbi/adbi/workflows/rcc/badge.svg)](https://github.com/r-dbi/adbi/actions)
[![Codecov test coverage](https://codecov.io/gh/r-dbi/adbi/graph/badge.svg)](https://app.codecov.io/gh/r-dbi/adbi/tree/main)
[![CRAN status](https://www.r-pkg.org/badges/version/adbi)](https://CRAN.R-project.org/package=adbi)
<!-- badges: end -->

Bringing [arrow-adbc](https://github.com/apache/arrow-adbc) to R via [DBI](https://github.com/r-dbi),
adbi aims to provide DBI-compliant database access.

## Goals and non-goals

adbi aims to:

- Expose ADBC drivers through DBI,
  so that a connection opened with `dbConnect(adbi("adbcsqlite"), ...)` behaves like any other DBI connection.
- Accept any driver built on the adbcdrivermanager framework,
  passed as an `adbc_driver` object, as a function returning one, or as a `pkg::fun` string.
- Support DBI's Arrow extension next to the `data.frame` API,
  so that results can stay in the Arrow format, fetched whole or chunk by chunk.
- Offer an efficient alternative to ODBC for analytical applications.
- Conform to the DBI specification, which DBItest checks against the adbcsqlite driver.

It is explicitly not trying to:

- Implement ADBC.
  The C API and the driver manager live in arrow-adbc and adbcdrivermanager; adbi is the DBI layer on top of them.
- Ship database drivers.
  Packages such as adbcsqlite are suggested rather than imported, and you install the ones you need.
- Define the database interface.
  DBI owns the generics and their specification.
- Hide what an individual ADBC driver cannot do.
  Temporary tables, 64-bit integers and identifier quoting follow the driver,
  and the DBItest setup under `tests/testthat/` records where that shows.

## Installation

Install the latest release of adbi from CRAN with the following code:

```r
install.packages("adbi")
```

To get a bug fix or to use a feature from the development version,
you can install the development version of adbi from GitHub:

``` r
# install.packages("devtools")
devtools::install_github("r-dbi/adbi")
```

## Example

The `data.frame` API of DBI is supported.

```{r df}
# To run this example, please install the adbcsqlite package first.
# 
# install.packages("adbcsqlite")

library(DBI)

# Create an SQLite connection using the adbcsqlite backend
con <- dbConnect(adbi::adbi("adbcsqlite"), uri = ":memory:")

# Write a table
dbWriteTable(con, "swiss", datasets::swiss)

# Query it
dbGetQuery(con, "SELECT * from swiss WHERE Agriculture < 40")

# Prepared statements
res <- dbSendQuery(con, "SELECT * from swiss WHERE Agriculture < ?")

dbBind(res, 30)
dbFetch(res)

dbBind(res, 20)
dbFetch(res)

# Cleanup
dbClearResult(res)
```

More interestingly, the recent arrow-extension API of DBI is supported as well.

```{r arrow}
# Queries
dbGetQueryArrow(con, "SELECT * from swiss WHERE Agriculture < 40")

# Prepared statements
res <- dbSendQueryArrow(con, "SELECT * from swiss WHERE Agriculture < ?")

dbBind(res, 30)

ret <- dbFetchArrow(res)
ret$length

dbBind(res, 20)

# Chunked fetches
while (!dbHasCompleted(res)) {
  ret <- dbFetchArrowChunk(res)
  message("fetched ", ret$length, " rows")
}

# Cleanup
dbClearResult(res)
dbDisconnect(con)
```
